---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r}
#Define the grid
#how many 'parts' we want to split up the p in, 0-1, as many as it takes to make a smooth distribution
dens = 20 
#creating the grid with p from 0 to 1 split in 20 (dens)
p.grid = seq(from=0, to=1, length.out=dens)

#Define the prior
#Flat (uniform) prior
prior = rep(1,dens)
#prior = ifelse(p.grid < 0.5, 0, 1) a more optimistic prior only considering p above chance

#Compute the likelihood at each value in grid
likelihood_RF = dbinom(3, size = 6, prob = p.grid)

#Compute the posterior (likelihood by prior)
unstd.posterior_RF = likelihood_RF*prior

#Standardize the posterior (so it sums to 1)
posterior_RF = unstd.posterior_RF/sum(unstd.posterior_RF)

#Put the values together in a dataframe
Data_RF=data.frame(grid=p.grid,posterior=posterior_RF,prior=prior,likelihood=likelihood_RF)

#Plot the posterior
posterior_plot_RF = ggplot(Data_RF,aes(grid,posterior))+geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+xlab("Knowledge of CogSci")+ylab("posterior probability")

#The probability that Riccardo knows more than chance (sum of the posterior where hypothesis is that he knows more than chance)
sum(Data$posterior[Data$grid>0.5])

#Quadratic approximation, instead of trying all like in grid approximation we assume that the distribution is gaussian and therefore we only need mean and SD to describe it, we test one, go in one direction - if it goes up continue if not go in the other direction until it starts to go down - we have found the mean. But if the shape assumption is not respected we get wrong results - this is also the case in frequentist statistics

library(rethinking)
globe.qa <- map(
    alist(
        w ~ dbinom(6,p) ,  # binomial likelihood
        p ~ dunif(0,1)     # uniform prior
), data=list(w=3) )
# display summary of quadratic approximation
precis( globe.qa )


```

2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.

```{r}
#Likelihood,posterior, standardized posterior, dataframe and plot KT
likelihood_KT = dbinom(2, size = 2, prob = p.grid)

unstd.posterior_KT = likelihood_KT*prior

posterior_KT = unstd.posterior_KT/sum(unstd.posterior_KT)

Data_KT=data.frame(grid=p.grid,posterior=posterior_KT,prior=prior,likelihood=likelihood_KT)

posterior_plot_KT = ggplot(Data_KT,aes(grid,posterior))+geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+xlab("Knowledge of CogSci")+ylab("posterior probability")


#Likelihood,posterior, standardized posterior, dataframe and plot JS
likelihood_JS = dbinom(160, size = 198, prob = p.grid)

unstd.posterior_JS = likelihood_JS*prior

posterior_JS = unstd.posterior_JS/sum(unstd.posterior_JS)

Data_JS=data.frame(grid=p.grid,posterior=posterior_JS,prior=prior,likelihood=likelihood_JS)

posterior_plot_JS = ggplot(Data_JS,aes(grid,posterior))+geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+xlab("Knowledge of CogSci")+ylab("posterior probability")

#Likelihood,posterior, standardized posterior, dataframe and plot MW
likelihood_MW = dbinom(66, size = 132, prob = p.grid)

unstd.posterior_MW = likelihood_MW*prior

posterior_MW = unstd.posterior_MW/sum(unstd.posterior_MW)

Data_MW=data.frame(grid=p.grid,posterior=posterior_MW,prior=prior,likelihood=likelihood_MW)

posterior_plot_MW = ggplot(Data_MW,aes(grid,posterior))+geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+xlab("Knowledge of CogSci")+ylab("posterior probability")


#Prior plot
prior_plot = ggplot(Data_RF,aes(grid,prior))+geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+xlab("Knowledge of CogSci")+ylab("posterior probability")

#Plot RF
posterior_plot_RF

#plot MW
posterior_plot_MW

#Both plots are centred around 0.5 (chance) however Riccardos is wider - probably because he has less observations so the posterior is less sure
```

3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

```{r}
dens_optimistic = 10000
prior_optimistic = rnorm(dens_optimistic,0.8,0.2)

#Optimistic posterior RF
unstd.posterior_RF_optimistic = likelihood_RF*prior_optimistic

posterior_RF_optimistic = unstd.posterior_RF_optimistic/sum(unstd.posterior_RF_optimistic)

Data_RF_optimistic=data.frame(grid=p.grid,posterior=posterior_RF_optimistic,prior=prior_optimistic,likelihood=likelihood_RF)

posterior_plot_RF_optimistic = ggplot(Data_RF_optimistic,aes(grid,posterior_RF_optimistic))+geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior_optimistic/dens_optimistic),color='red')+xlab("Knowledge of CogSci")+ylab("posterior probability")


```

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r}



```

5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

```{r}



```

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

```{r}



```

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



